\begin{section}{Monomer-Based Parameterization}
\label{sec:workflow-monomer_parameters}

While/after running the \dftsapt calculations, the next step in the Workflow
is to compute various force field parameters which only depend on the
identities of the individual monomers themselves. The following subsections describe the
calculations of multipole moments (\cref{sec:workflow-multipoles}),
short-range exponents (\cref{sec:workflow-exponents}), dispersion coefficients
(\cref{sec:workflow-dispersion}), and induced dipole polarizabilities
(\cref{sec:workflow-polarizabilities}). First, however, we 
outline the scope and useful features of the \camcasp software used to
perform these monomer property calculations.

\begin{subsection}{Distributed Property Calculations using \camcasp}

\camcasp is a collection of scripts and programs useful for (among other
things) the calculation of distributed multipoles and
polarizabilities.\cite{camcasp5.8} Of particular importance is the
choice of distribution method, as this determines how the various molecular properties
of interest should be mapped onto corresponding atom-in-molecule properties. Currently, two
main distribution (or `charge partitioning') schemes are available in \camcasp:
\dma\cite{Stone2005} and
\isa.\cite{Misquitta2014} The theory behind the \isa procedure has already been detailed in
\cref{ch:introduction}, and monomer property calculations using \dma are
described in 
\citen{Stone2005,Misquitta2006,McDaniel2014a}. In general, and where
available, \isa-based properties are to be preferred, and we recommend an
\isa-based parameterization scheme for obtaining multipoles and
atom-in-molecule exponents. A \dma-based method is currently required for
obtaining dispersion coefficients and static polarizabilities, though
\isa-based strategies for these properties are under active development and
(in the case of dispersion) are discussed in \cref{sec:workflow-alston}.
A complete overview of
available property calculations and distribution schemes, along with relevant references,
is given in \cref{tab:workflow-distribution_schemes}.


\begin{table}[ht]
\centering
\begin{tabular}{@{}lcc@{}}
\hline
\toprule
\multirow{2}{*}{Property}
& \multicolumn{2}{c}{Parameterization Scheme} \\
\cmidrule{2-3} 
                         &  ISA                                    & DMA          \\ 
\midrule
\multirow{2}{*}{Multipoles}               & \cref{sec:workflow-multipoles}          &                      --               \\  
                                          & \citen{Misquitta2014}                   & \citen{Stone2005,McDaniel2014a}                 \\  
\addlinespace
\multirow{2}{*}{Exponents}                &  \cref{sec:workflow-exponents}          & \multirow{2}{*}{--}                   \\ 
                                          &  \citen{VanVleet2016}                   &                                       \\ 
\addlinespace
\multirow{2}{*}{Dispersion Coefficients}  &  \cref{sec:workflow-dispersion}         & \cref{sec:workflow-dispersion}        \\ 
                                          &  --                                     & \citen{McDaniel2013}        \\ 
\addlinespace
\multirow{2}{*}{Dipole Polarizabilities}  &  \multirow{2}{*}{--}                    & \cref{sec:workflow-polarizabilities}  \\ 
                                          &                                         & \citen{McDaniel2013}                  \\ 
\addlinespace
\bottomrule
\hline
\end{tabular}
\caption
[Overview of \isa- and \dma-based methods for obtaining distributed monomer
properties]
{Overview of \isa- and \dma-based methods for obtaining distributed monomer
properties. Details for each monomer parameterization are given in the listed
section and/or reference.}
\label{tab:workflow-distribution_schemes}
\end{table}

\end{subsection}



%%%%%%%%%%%%%%%% MULTIPOLES %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsection}{Multipoles}
\label{sec:workflow-multipoles}

\begin{subsubsection}{Practicals}

\isa-based multipoles are described in detail in \citen{Misquitta2014}, and can
be calculated using the \camcasp software. To set-up the \isa calculations,
execute the command
%
\begin{lstlisting}
./scripts/make_isa_files.py
\end{lstlisting}
%
which creates the necessary \isa files for calculating both distributed
multipoles and exponents (see \cref{sec:workflow-exponents}). After running
these calculations (a process which may require several hours, depending on
the molecule), the multipoles can be extracted simply by running
%
\begin{lstlisting}
./scripts/workup_isa_charges.py
\end{lstlisting}
%
This work-up script produces several output files,
\begin{itemize}[noitemsep,label=]
\item \verb|<monomer>_ISA_L4.mom|
\item \verb|<monomer>_ISA_L2.mom|
\item \verb|<monomer>_ISA_L0.mom|
%\item \verb|<monomer>_ISA_L4_to_L0.mom|
\end{itemize}
which correspond to multipole moments for various long-range electrostatic
models. Using Stone's notation,\cite{stone2013theory} the $Lx$ suffix refers
to the highest order of multipole moments ($L0$ = point charges, $L1$ =
dipoles, $L2$ = quadrupoles, etc.) included in the model. The $L4$ model is
output by the \camcasp software package, and the $L2$ and $L0$ models are
generated by rank-truncation (that is, zeroing out) of the higher-order multipole
moments. 
For most routine
force field development, the $L2$ model is to be preferred for its balance of
accuracy and computational expense. Next, however, we
discuss situations in which different electrostatic models may be desirable.

\end{subsubsection}
\begin{subsubsection}{Advanced Multipole Parameterization Options}

As stated above, for the purposes of obtaining sub-\kjmol{} accuracy
force fields it is often important to model the long-range electrostatics
using \isa-based multipoles truncated to no farther than quadrupolar (i.e. `rank 2' or
$L2$)\cite{stone2013theory} contributions.
Due to computational and/or software limitations, however, there exist
practical cases where it becomes advantageous to exclude all higher-order
multipole moments.\cite{Cardamone2014} In such cases, 
two different types of long-range electrostatic models are useful. First, for
reasonably isotropic molecules a good option is to rank-truncate the \isa
multipoles to the $L0$ point charge contributions, thus yielding a so-called `atom-centered
point charge model'. On the other hand, for more anisotropic functional groups such as those
described in \citen{Kramer2014}, an atom-centered point charge model can be
insufficiently accurate, making it 
necessary to model the long-range electrostatics by including additional 
`off-center/off-site' point charges. Given a well-chosen set of off-site charges, 
an off-center point charge model usually can reasonably reproduce the effects of the
neglected higher-order multipole moments.\cite{Dixon1997}
%% (Using water as an example, off-site charges have been shown to substantially improve
%% force field accuracy, and can either be placed in a
%% tetrahedral arrangement, requiring two additional charge sites, or in the
%% \ch{H-O-H} plane itself, requiring one additional charge site.\cite{Tran2016}) 
In the past, locations for the off-center charges have usually been manually tuned or optimized
in a system-specific manner, though recent work suggests the
possibility of switching to non-empirical methods in order to more easily calculate/optimize 
positions for the extra-atom sites.\cite{Chaudret2013,Unke2017} 

For atom-centered point charge models, the output of the
\verb|workup_isa_charges.py| script automatically provides the required
rank-truncated multipole file (listed as \verb|<monomer>_ISA_L0.mom| in the \verb|isa/|
sub-directory). Note that, because the \verb|<monomer>_ISA_L0.mom| file is given as a
simple rank-truncation of the more complete \verb|<monomer>_ISA_L2.mom|
multipoles, the $L0$ moments (that is, point charges) are identical between the two files.

For developing rank-transformed point charge models, \citeauthor{Ferenczy1997} has
programmed a method for calculating electrostatic potential-fitted charges,
which can be thought of as a `rank transformation' procedure. The author's
MULFIT program can be 
downloaded online at
\url{http://www-stone.ch.cam.ac.uk/pub/gdma/index.php}, and documentation for
the program is found in the \verb|documentation/| sub-directory of the
Workflow. Assuming the \verb|mulfit| executable is in your \verb|$PATH|, a
basic rank transformation can be performed using the following steps:
%
\begin{lstlisting}[language=bash]
cp templates/mulfit.inp isa/<monomer>/OUT/
cd isa/<monomer>/OUT/
mulfit < mulfit.inp
\end{lstlisting}
%
Here the default \verb|mulfit.inp| file is set to take in the $L4$ rank
multipoles and rank-transform them to an $L0$ model. In this case, note that
the $L0$ moments between the rank-transformed and rank-truncated moments will
\emph{not} be identical, and testing is required to ascertain which
moments yield optimal force field parameters.

The MULFIT program can additionally be used to develop off-site point charge models. In
this case, the input multipole file (default \verb|ISA_L4.mom|) should be
edited to include the additional sites, and an example of the required syntax
is given in \verb|documentation/examples/ISA_L4_offsites.mom| for a
4-site water model. Importantly, the MULFIT program does not help
optimize the position(s) of the off-site charge(s), and thus the task of choosing
the number and position(s) of the off-site(s) is left to
the user.

After fitting multipole parameters with the MULFIT program, the program output
gives two indications of fit quality. First, the agreement between the total
reference and fitted multipoles moments is listed, and this should be taken as
a primary indication of multipole quality. Second, the program gives a
`Goodness of fit' parameter, expressed as an energy. While difficult to
interpret in an absolute sense, in comparing different rank-transformed models
we have generally found that models with lower `Goodness of fit' parameters
yield better force field fits.

\end{subsubsection}
\end{subsection}
%%%%%%%%%%%%%%%% MULTIPOLES %%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%% EXPONENTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsection}{ISA Exponents}
\label{sec:workflow-exponents}

%\begin{subsubsection}{Overview}

As described in \cref{ch:introduction,ch:isaff}, the \isa procedure
produces a set of distributed \aim electron densities. The
orientational average of each of these \aim densities, or
`shape-functions', are spherically-symmetric quantities that describe the
radial decay of the \aim density.\cite{Misquitta2014} As described in
\cref{ch:isaff}, and using the algorithm detailed in 
\cref{sec:workflow-exponent_algorithm},
the shape-functions can be fit to a Slater-type function in
order to yield an isotropic, exponentially-decaying model for the \isa
densities. Importantly, the Slater-exponents in this density model directly
yield the exponents necessary to describe short-range effects (such as
exchange-repulsion and charge penetration) in the two-body force
field (see \cref{ch:isaff} for details).

Assuming the \isa calculations have already been run to obtain multipole
moments (see previous section), the \isa exponents can be obtained very simply
by running the command
%
\begin{lstlisting}
./scripts/workup_isa_exponents.py
\end{lstlisting}
%
The resulting exponents are given in the file \verb|isa/<monomer>.exp|, which
uses a file format recognized by the \pointer pre-prossessing scripts (see
\cref{ch:pointer}).

\end{subsection}
%%%%%%%%%%%%%%%% EXPONENTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%% DISPERSION %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsection}{Dispersion Coefficients}
\label{sec:workflow-dispersion}

\input{workflow/dispersion.tex}

\end{subsection}
%%%%%%%%%%%%%%%% DISPERSION %%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%% POLARIZATION %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{subsection}{Polarization Charges}
\label{sec:workflow-polarizabilities}

\input{workflow/induction.tex}

\end{subsection}
%%%%%%%%%%%%%%%% POLARIZATION %%%%%%%%%%%%%%%%%%%%%%%%%

\end{section}
