%% \begin{wbepi}{David C.~Makinson (1965)}
%% It is customary for authors of academic books to include in their prefaces statements such as this: ``I am indebted to ... for their invaluable help; however, any errors which remain are my sole responsibility.'' Occasionally an author will go further. Rather than say that if there are any mistakes then he is responsible for them, he will say that there will inevitably be some mistakes and he is responsible for them....
%% 
%% Although the shouldering of all responsibility is usually a social ritual, the admission that errors exist is not --- it is often a sincere avowal of belief. But this appears to present a living and everyday example of a situation which philosophers have commonly dismissed as absurd; that it is sometimes rational to hold logically incompatible beliefs.
%% \end{wbepi}

%% Above is the famous ``preface paradox,'' which illustrates how to use the \texttt{wbepi} environment for epigraphs at the beginning of chapters.  You probably also want to thank the Academy.

From my time as a graduate student, I owe an unspeakable debt of gratitude to
the people who have been and continue to be in my life. Not everyone is
mentioned here with the attention that they deserve, in part because the list
would be far too long, and in part because the meaning you've added to my life
is impossible to express in a few short sentences. Nevertheless, know that
your impact is felt daily, and continues to make this life worthwhile.

To my undergraduate research mentors, Bill, Dave, Nancy, and Bob: Every day in grad
school that I doubted, your voices kept me believing in myself and my
abilities. 

To my graduate research mentor, Jesse: I could not have asked for a more
patient and giving mentor. For all the time you gave me as a first year, for
the many questions I asked and you answered, and for the frequent conversations
that gave me insight and direction into my project, I owe you a large debt of
gratitude. Your students at Georgia Tech are going to be incredibly lucky to
have you on their side.

To my current and former group members, Kuang, Jesse, Jeff, Glen, Eric,
Benjamin, Tingting, Chenyang, Aurora, Xinyi, Nina, and Tesia: In all my five
years of graduate school, I found every single one of our interactions to be
rewarding, thought-provoking, and just plain fun. From those of whom I required much
of your time (Eric and Jesse especially), thank you for your kindness in
growing me into a better scientist. For those of you whom I mentored (Tesia in
particular), thank you for your patience and enthusiasm as I haltingly learned
how to instruct.  I wish you all the very best in the future.

To Alston: JR and I shouldn't have made nearly as much progress in
force field development had it not been for our timely and rewarding
collaboration. Thank you for your enthusiasm for life and science, and for all
that you've taught me since our first interactions.

To the UW-Madison faculty and staff, especially Clark, Desiree, Alan, Cheri, Randy,
Arietta, Amanda, Qiang, Arun, and Ned: In interactions big and small, you made me feel
that I belong at this institution. Clark in particular, thank you for
your constant affirmation of my career goals; hearing this verbally from a
committee member has meant a lot to me.

To my advisor, JR: There are few scientists whom I admire and respect as much
as you. (I would say `no' scientists, but you've always cautioned me
against using extreme language in scientific communication.) Your integrity,
humility, and passion for science have enabled me to be a far better student
than would have been possible otherwise, and I can only hope to learn how to
mentor others with the same thoughtfulness and care that you gave me.

To my friends at Madison, past and present (Britta, Diane, Bob and Deb, Cale, Ryan V.,
Stephanie, Aaron, Ryan R., Adam, Nick, and all the rest): The relationships
I've had here will make me miss Wisconsin, which is something I couldn't have
imagined saying about a state without mountains. Britta, thank you for feeding
me while I wrote this thesis, and for not complaining about the messes I made.
Diane, thank you for trusting me and teaching me frisbee. Stephanie, thank you
for reminding me how to be passionate about God and social justice.  Cale and
Ryan V., thank you for welcoming me into your lives when I was new here.
Nick, thank you for making me feel understood.  Aaron, thank you for
everything.

To my family, both blood relations and Narwhals: I've needed you all in this
graduate school journey, at times to make me feel safe, and at times to
challenge me both personally and academically. Thank you for your support, and
thank you for your love. It's been the deepest honor to have you all in my
life, and I look forward to many more years of journeying together.

To my partner, Alex: I truly couldn't have asked for a more amazing person to share my
life with. Thank you for standing beside me in times of stress and sorrow,
thank you for taking me out to tea and laughing with me in times of happiness,
and thank you for giving me the opportunity to love you and be loved in return.

To God: Thank you for making this universe so full of beauty and mystery.


